---
title: "methy_analysis"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(knitr)
library(limma)
library(RColorBrewer)
library(stringr)
library(data.table)

```


```{r  methylation}

TCGA.Metadata = read.table("BRCA_clinicalMatrix_New.txt", header = TRUE, check.names = FALSE, sep = "\t", row.names = NULL)

ann450k = getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)
ann450k = data.frame(ann450k)

#Methylation props vs bulk? 
methy = read.table("HumanMethylation450", header = TRUE)
rownames(methy) = methy$sample
methy$sample = NULL
methy = na.omit(methy)
colnames(methy) = gsub("\\.","-",colnames(methy) )

#Need props 
tcga_ggtern = read.table("props.txt")
tcga_ggtern_ep = tcga_ggtern[,c(1,7,9)]
tcga_ggtern_ep[tcga_ggtern_ep< 0] <-0

tcga_ggtern_ep$Colsum = tcga_ggtern_ep$Epithelial.Basal + tcga_ggtern_ep$Epithelial.Her2 + tcga_ggtern_ep$Epithelial.Luminal
tcga_ggtern_ep$Epithelial.Her2 = tcga_ggtern_ep$Epithelial.Her2 * (1/tcga_ggtern_ep$Colsum)
tcga_ggtern_ep$Epithelial.Basal = tcga_ggtern_ep$Epithelial.Basal * (1/tcga_ggtern_ep$Colsum)
tcga_ggtern_ep$Epithelial.Luminal = tcga_ggtern_ep$Epithelial.Luminal * (1/tcga_ggtern_ep$Colsum)
tcga_ggtern_ep = na.omit(tcga_ggtern_ep)


#Bulk methylation

genes_of_interest = c("GRB7" , "MMP7", "GSTP1")

result <- filter(ann450k, grepl(paste(genes_of_interest, collapse="|"), UCSC_RefGene_Name))

all_matches = ann450k$UCSC_RefGene_Name[grepl(paste(genes_of_interest, collapse = "|"), ann450k$UCSC_RefGene_Name)]

genes_anno = ann450k[ann450k$UCSC_RefGene_Name %in% all_matches,]
genes_anno = genes_anno[!genes_anno$UCSC_RefGene_Name  == "RASSF10",]

probes_interest = genes_anno$Name

methy_probes_interest = methy[rownames(methy) %in% probes_interest,]
methy_probes_interest = data.frame(t(methy_probes_interest))

#Make sure that we have the ggten propos for these 
setdiff(rownames(methy_probes_interest),tcga_ggtern$Sample)
methy_probes_interest = methy_probes_interest[rownames(methy_probes_interest) %in% rownames(tcga_ggtern_ep),]


#Top quantile methylated for each genes 

top_methy_sample_TGCA = list()
for(i in 1:length(genes_of_interest)){
  probes = genes_anno[genes_anno$UCSC_RefGene_Name %like% genes_of_interest[[i]], "Name" ]
  probes_df = methy_probes_interest[,colnames(methy_probes_interest) %in% probes]
  probes_df$means = rowMeans(probes_df)
  quantile(probes_df$means)
  top = as.numeric(quantile(probes_df$means)[4])
  probes_df = probes_df[probes_df$means >= top,]
  top_methy_sample = rownames(probes_df)
  top_methy_sample_TGCA[[genes_of_interest[[i]]]] = top_methy_sample
}


#top25
top_methy_sample_TGCA_25 = list()
for(i in 1:length(genes_of_interest)){
  probes = genes_anno[genes_anno$UCSC_RefGene_Name %like% genes_of_interest[[i]], "Name" ]
  probes_df = methy_probes_interest[,colnames(methy_probes_interest) %in% probes]
  probes_df$means = rowMeans(probes_df)
  probes_df = probes_df[order(-probes_df$means),]
  probes_df = probes_df[1:25,]
  top_methy_sample = rownames(probes_df)
  top_methy_sample_TGCA_25[[genes_of_interest[[i]]]] = top_methy_sample
}
```

```{r ggtern}

###FIGURE 6##

plots_ggtern_methy = list()
for(i in 1:length(top_methy_sample_TGCA_25)){
  gene = names(top_methy_sample_TGCA_25)[[i]]
  samples_methylated = top_methy_sample_TGCA_25[[i]]
  tcga_ggtern_gene =  tcga_ggtern_ep[rownames(tcga_ggtern_ep) %in% samples_methylated,]
 plots_ggtern_methy[[gene]] = ggtern(tcga_ggtern_gene, aes(x = Epithelial.Luminal,y = Epithelial.Basal, z = Epithelial.Her2))  + geom_point(color="black",size=1)  
}

for(i in 1:length(plots_ggtern_methy)){
  pdf(paste(names(plots_ggtern_methy)[i],"point_methy.pdf",sep = "."))
  plot(plots_ggtern_methy[[i]])
  dev.off()
  }


tcga_ggtern_ep

table_pathway_mapk = table_pathway[,c(1,7,16:18,20)]
rownames(table_pathway_mapk) = table_pathway_mapk$name
table_pathway_mapk$name = NULL
table_pathway_mapk = table_pathway_mapk[order(-table_pathway_mapk$MAPK),]
table_pathway_mapk = table_pathway_mapk[1:25,]

pdf("mapk.pdf")
ggtern(table_pathway_mapk, aes(x = Epithelial.Luminal,y = Epithelial.Basal, z = Epithelial.Her2)) +
stat_density_tern(geom='polygon',aes(fill=..level..,
                             alpha = ..level..),n = 50, bins=10,color='grey',bdl = 0,contour = TRUE,base = "identity") + scale_fill_gradient(low = "blue",high = "red") + geom_mask() + geom_point(color="black",size=0.5)  
dev.off()




tcga_ggtern_all_x = tcga_ggtern_ep[top_methy_sample_TGCA$MMP7,]
ggtern(tcga_ggtern_all_x, aes(x = Epithelial.Luminal,y = Epithelial.Basal, z = Epithelial.Her2)) +
stat_density_tern(geom='polygon',aes(fill=..level..,
                             alpha = ..level..),n = 400, bins=10,color='grey',bdl = 0,contour = TRUE,base = "identity") + scale_fill_gradient(low = "blue",high = "red") + geom_mask() + geom_point(color="black",size=0.5)  
